---
title: "R Project"
author: "Adel"
date: "12/19/2020"
output: html_document
---
Download nessary packages/libraries
```{r message=FALSE, warning=FALSE}
library("ggplot2")
library("ggthemes")
library("dplyr")
```
Download all data sets of the three cities.

```{r}
ny = read.csv('C:\\Users\\adela\\OneDrive\\Documents\\R DIr\\new-york-city.csv')
chi = read.csv('C:\\Users\\adela\\OneDrive\\Documents\\R DIr\\chicago.csv')
wash = read.csv('C:\\Users\\adela\\OneDrive\\Documents\\R DIr\\washington.csv')
```
Exploring data for each data set
```{r}
head(ny)
head(chi)
head(wash)
```
Q1: Does Gender influence the type of rider in New York?




Running a descriptive summary to understanding the data
```{r}
summary(ny)
```
The summary showed that both Gender and User Type  have blank cells which seem to be missing values.Ploting User type by Gender in New York using barcharts:
```{r}
ggplot(data= ny, aes(x= Gender, fill = User.Type))+geom_bar(position = 'stack') + ylab('Number of riders')+ggtitle('Fig 1:User type by Gender in NY')

```


As the graph confirms our findings above with regards to variables of interest,
Choice is made to consider the blank cells as NA's.

```{r message=FALSE, warning=FALSE}
ny$Gender[ny$Gender==""] <-NA
ny$User.Type[ny$User.Type ==""] <-NA
```
Now , Let's plot gain the distribution of User type by gender without missing values (excluding NA's) to get a more clearer picture.
```{r}
ny %>%
  filter(!is.na(Gender) & !is.na(User.Type)) %>%
  ggplot(aes(x = Gender, fill = User.Type))+
  geom_bar(position = "stack") + 
  ylab('Number of riders') +
  ggtitle("Fig 2: User type by Gender in NY without missing values")
```


Running a compute a summary table to represent the total counts of customers and suscribers according to their gender.
```{r}
ny2 <-ny %>%
  filter(!is.na(Gender) & !is.na(User.Type))

by( ny2$User.Type ,ny2$Gender , summary)
```
Performing a chi-square statistic test to establish wether Gender and user type are asociated?
#H0: Gender and user type are independent
#H1: Gender and usere type are related

```{r}
chisq.test(ny2$Gender, ny2$User.Type, correct =FALSE)
```
since the P-value is less than 0.05 , we can conclude that there is a relationship between Gender and User Type.



Q2:What are the counts of each gender (only available for NYC and Chicago)?

```{r message=FALSE}
chi %>% group_by(Gender) %>% filter(!is.na(Gender))%>%
  summarise(total = length(Gender))%>%
  ggplot(aes(Gender, total)) + geom_bar(stat = 'identity')+
  xlab("Gender") +  ylab("Count ")+ ggtitle("Counts of each gender in chicago")


ny %>% group_by(Gender) %>% filter(!is.na(Gender))%>%
  summarise(total = length(Gender))%>%
  ggplot(aes(Gender, total)) + geom_bar(stat = 'identity')+
  xlab("Gender") +  ylab("Count ")+ ggtitle("Counts of each gender in new york")
```



Females are lesser in number as compared to Males in chicago and newyork.observations where gender were NA were filtered out




Q3:What is the gender wise median travel time for users in different cities ?

```{r}
summary(chi$Trip.Duration)
```
Median trip duration in chicago is 670.0 seconds and Mean trip duration is 936.2 seconds.

```{r warning=FALSE}
ggplot(data = subset(chi, !is.na(Gender)), aes("Gender",Trip.Duration, fill = Gender)) + geom_boxplot()+  
  scale_y_continuous(limits = c(0,2000))+
  xlab("Gender") +  ylab("Time in Seconds ")+ ggtitle("Trip duration in Chicago")
```

```{r}
summary(ny$Trip.Duration)
```

Median trip duration in newyork is 609.0 seconds and Mean trip duration is 899.7 seconds.

```{r warning=FALSE}

ggplot(data = subset(ny, !is.na(Gender)), aes("Gender",Trip.Duration, fill = Gender)) + geom_boxplot()+  
  scale_y_continuous(limits = c(0,2000))+
  xlab("Gender") +  ylab("Time in Seconds ")+ ggtitle("Trip duration in NYC")
```
